---
title: "Introduction to prodigenr"
author: "Luke W. Johnston"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to prodigenr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Are you an academic researcher who often writes up abstracts for conferences or
submits manuscripts to journals? Do you often have to make slides or posters for
presentations? Is your usual workflow to copy a previous project and start 
replacing the old text for the new text? This R package was designed with you in
mind!

`prodigenr`, or *pro*ject *di*rectory *gen*erator, simplifies the process of
creating these new projects and can help make your workflow more reproducible.
Standard files and folders are created for specific projects (e.g. abstracts or
manuscripts), along with a workflow that tries to be simple and easy to use.

Because researchers often write or create many papers, slides, posters, and
abstracts, it can quickly become tedious and messy to always make a new
directory with all the necessary files and organization. 

## The main command: `prodigen`

To use `prodigenr`, you simply need to use the `prodigen` command. At present,
there are only four template projects that you can view using:

```{r templates, collapse=TRUE}
library(prodigenr)
list_templates()
```

These templates are projects that an academic researcher typically encounters.
However, if you have a suggestion or want to add a template, please create a
[Github issue](https://github.com/lwjohnst86/prodigenr/issues) or submit a [Pull Request](https://github.com/lwjohnst86/prodigenr/pulls)!

Starting a manuscript? Create a project directory like so:

```{r, eval=FALSE}
prodigen('manuscript', 'ManuscriptName', '~/path/to/new/project')
```

Do you use [Git](https://git-scm.com/)? You can start a git repository in the new project simply through the `git.init` argument:

```{r, eval=FALSE}
prodigen('manuscript', 'ManuscriptName', '~/path/to/new/project', git.init = TRUE)
```

Or if you need to use [Packrat](https://rstudio.github.io/packrat/), start it with `packrat.init`:

```{r, eval=FALSE}
prodigen('manuscript', 'ManuscriptName', '~/path/to/new/project', packrat.init = TRUE)
```

The same procedure is used for making the other project templates.

```{r, eval=FALSE}
prodigen('slides', 'PresentationName', '~/path')
prodigen('abstract', 'Name', '.') # Current directory
```

A `README.md` file is contained within each project that explains more about
what each folder does and what some of the files do that were created.

The end goal of each project is to be as self contained as possible. So that if
you ever need to go back to the analysis, it is easy to re-run the code and get
the results that you say you got. This is especially useful if others such as
reviewers ask for something or want to confirm your results. For more
information on good practices to use in making an analysis project, see 
[here](http://stats.stackexchange.com/questions/2910/how-to-efficiently-manage-a-statistical-analysis-project)
or 
[here](http://www.r-bloggers.com/managing-a-statistical-analysis-project-%E2%80%93-guidelines-and-best-practices/))

## Workflow when using projects created from `prodigenr`

A typical workflow would be to:

1. Write up your analysis and associated written explanations of the results, as
you would for any research project, in the abstract, poster, slides, or
manuscript `.Rmd` ([R Markdown](http://rmarkdown.rstudio.com/)) file.
2. Any piece of code you use more than once or is fairly complex, convert it
into a function. Put this new function into the `R/functions.R` file.
3. Fetch and wrangle your data in the `R/fetch_data.R` and to use the data, load
it using `load_data()`.
4. Knit the `.Rmd` file. You now have your final abstract, poster, slides, or
manuscript to use for your research.

## Related packages or projects

There are several ways of handling a project. There is at least two packages
that have similar goals as `prodigenr` as well as through the use of the R
package structure:

- [`ProjectTemplate`](http://projecttemplate.net/) is well documented and still
actively developed. Only downside is that it is fairly complicated to use and
complex in the project workflow it creates.
- [`makeProject`](https://cran.r-project.org/web/packages/makeProject/index.html)
is very simple and stripped down in what it creates and in it's use. Downside is
that it wasn't updated since 2012.
- Use of the R package structure via
[`devtools`](https://cran.r-project.org/web/packages/devtools/README.html),
which is argued for in this
[blog](https://rmflight.github.io/posts/2014/07/vignetteAnalysis.html) and
compared to `ProjectTemplate` in this
[blog](https://rmflight.github.io/posts/2014/07/zpackages_vs_projectTemplate.html)).

It's up to you to decide which style to use. The goal behind creating 
`prodigenr` was to incorporate the ideas behind R packages and mesh well with 
RStudio, while also trying to be as simple as possible so it is in some ways a
hybrid between `ProjectTemplate` and `makeProject`. It can always be improved, 
so if you have suggestions just submit a 
[GitHub issue](https://github.com/lwjohnst86/prodigenr/issues)!
